Linear Mixed Models: Random Intercept

### 2024-03-17 examples/_queue/1b_Examples_DifferenceScores_SimpleRegression.qmd
examples 1-4 updates

---
title: "Difference Scores: Simple Regression"
author: "Biostatistics Working Group"

reader-mode: true
---

@@

library(tidyverse)    # Collection of R packages for data science
library(easystats)    # Collection of R packages statistical modeling, visualization, and reporting
library(gtsummary)    # Publication-ready tables
library(report)       # Easy reporting of regression analyses
library(broom)        # Tidy and augment statistical models output
library(gridExtra)    # Arrange multiple grid-based plots on a page

>>>>

---
title: "Difference Scores: Simple Regression"
author: "Biostatistics Working Group"
theme: cosmo
---

@@

library(tidyverse)    # Collection of R packages for data science
library(easystats)    # Collection of R packages statistical modeling, visualization, and reporting
library(rstatix)      # Pipe-friendly framework for performing common stats
library(gtsummary)    # Publication-ready tables
library(report)       # Easy reporting of regression analyses
library(broom)        # Tidy and augment statistical models output
library(gridExtra)    # Arrange multiple grid-based plots on a page

### 2024-03-17 examples/_queue/3a_Examples_LinearMixedModels.qmd
examples 1-4 updates

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
:::

@@

```
#:::

@@

::: blue
**This code creates a descriptives table**
::: panel-tabset
### Read and View Data {.tabset .tabset-fade .tabset-pills}

:::
```{r descriptives}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

descriptives_table
```

## Results 
<div class = "blue">

@@

The code fits a linear mixed model to predict the 'Height' variable based on time points ('eventname') and handedness ('Handedness'), while accounting for individual-level variability by including random intercepts for each participant ('src_subject_id'). The results of the model are then printed to provide a summary of the fitted model parameters.

@@

## Output and reports extending from the LMM-ri analyses
#summary(model)
#confint(model, level = 0.95, method = "Wald")
model_performance(model)
model_parameters(model)
model_parameters(model,standardize = "refit")

@@

The following set of plots are used to facilitate model diagnostics. The first is a histogram showcasing the distribution of random intercepts for individual subjects, indicating variations in height not explained by the fixed effects. The second depicts residuals versus fitted values, helping assess the model's fit and potential heteroscedasticity. The third contrasts observed and predicted height values across different time points, offering a side-by-side evaluation of the model's predictions against actual observations.

::: blue
```{r diagnosticplots}

@@

```{r}
plot(parameters(model)) +
  ggplot2::labs(title = "A Dot-and-Whisker Plot")

plot(model)
theme_set(theme_modern())
result <- estimate_expectation(model, data = "grid")

plot(result)

plot(result, type = "qq")
plot(result, type = "pp")

normality <- check_normality(model, effects = "random")
plot(normality)

heteroscedasticity <- check_heteroscedasticity(model)
plot(heteroscedasticity)

homogeneity <- check_homogeneity(model)
plot(homogeneity)
plot(homogeneity, type = "density")

check_model(model)

check_model(model, panel = FALSE)

distribution <- check_distribution(model)
distribution
plot(distribution)

contrasts <- estimate_contrasts(model, contrast = "sex")
means <- estimate_means(model)
plot(means)
plot(contrasts, means)

params <- model_parameters(model, effects = "fixed")
plot(params)

plot(params, show_labels = TRUE, size_text = 4)

@@

```{r}
report(model)
```

>>>>

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
:::
:::

## Descriptives Overview

::: panel-tabset
### Read and View Data {.tabset .tabset-fade .tabset-pills}

@@

@@

::: blue
**This code creates a descriptives table**

```{r descriptives}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

@@

descriptives_table
```
:::
:::

## Results

@@

::: blue
The code fits a linear mixed model to predict the 'Height' variable based on time points ('eventname') and handedness ('Handedness'), while accounting for individual-level variability by including random intercepts for each participant ('src_subject_id'). The results of the model are then printed to provide a summary of the fitted model parameters.

@@

## Output and reports extending from the LMM-ri analyses
#summary(model)
#confint(model, level = 0.95, method = "Wald")

model_performance(model)

model_parameters(model)

model_parameters(model,standardize = "refit")

@@

::: blue
The following set of plots are used to facilitate model diagnostics. The first is a histogram showcasing the distribution of random intercepts for individual subjects, indicating variations in height not explained by the fixed effects. The second depicts residuals versus fitted values, helping assess the model's fit and potential heteroscedasticity. The third contrasts observed and predicted height values across different time points, offering a side-by-side evaluation of the model's predictions against actual observations.

```{r diagnosticplots}

@@

```{r}

library(ggplot2)
theme_set(theme_modern(base_size = 6))

plot(parameters(model)) +
  ggplot2::labs(title = "A Dot-and-Whisker Plot")
#plot(model)

result <- estimate_expectation(model, data = "grid")
plot(result)
#plot(result, type = "qq")
#plot(result, type = "pp")

normality <- check_normality(model, effects = "random")
#plot(normality)

heteroscedasticity <- check_heteroscedasticity(model)
#plot(heteroscedasticity)

homogeneity <- check_homogeneity(model)
#plot(homogeneity, type = "density")

distribution <- check_distribution(model)
#plot(distribution)

#contrasts <- estimate_contrasts(model, contrast = "sex")
#means <- estimate_means(model)
#plot(means)
#plot(contrasts, means)

params <- model_parameters(model, effects = "fixed")
#plot(params, show_labels = TRUE, size_text = 4)

#check_model(model)

@@

```{r}
report(model)
```
:::
